Topological Eulerian Synthesis of Slow Motion Periodic Videos

### 2018-02-04 Paper/paper.tex
Adding GUDHI reference, changing batch tests to go through Alpha instead of Kappa
probably conflicts

\graphicspath{{Figures/}}
\newcommand{\argmin}{\operatornamewithlimits{argmin}}
\newcommand{\argmax}{\operatornamewithlimits{argmax}}
\newcommand{\matt}[1]{{\color{purple}[MSB: #1]}}

@@

\textbf{Preprocessing.} The input to our method is a 3-channel video of $N$ frames, each frame of resolution $W \times H$. We treat each frame as a point in Euclidean space of dimension $\mathbb{R}^{3 \cdot W \cdot H}$, and denote the set of all videos by the sequence $\mathbf{X}(t) \in \mathbb{R}^{(3 \cdot W \cdot H)}$ for integer-valued $t \in [1,N]$. The steps of our approach outlined in Figure~\ref{fig:ConceptFigure}(a-c), however, do not rely on the {\em extrinsic} geometry of the set of points $\mathbf{X}(t)$, only its set of {\em pairwise distances}. Thus, we can project all of the points into a lower-dimensional space and apply the set of techniques to the projection, assuming that pairwise distances are preserved. Since $N \ll (3 \cdot W \cdot H)$, the frames can be projected into a $N-1$-dimensional space such that all distances are preserved. Denoting this projection by $\mathbf{P} \in \mathbb{R}^{(3 \cdot W \cdot H) \times (N-1)}$, our method for steps (a)-(c) operates on $\mathbf{Z}(t) = \mathbf{P} \mathbf{X}(t)$.

@@

Given a $W \times H$ video $X(t) \in \mathbb{R}^{3WH}$ indexed by time, the sliding window video embedding \cite{cao1998dynamics,traliehigh,tralie2017quasi} is defined as

\begin{equation}
SW_{\text{dim}, \tau}[X(t)] = \left[ \begin{array}{c} X(t) \\ X(t + \tau) \\ \vdots \\ X(t + (\text{dim}-1)\tau)  \end{array} \right] \in \mathbb{R}^{3WH(\text{dim})}
\end{equation}

where $\tau$ is a step size and ``dim'' is the dimension of the embedding.  In this work, we fix $\tau = 1$ for simplicity and drop $\tau$ in the subscript, and we also sample times $t$ only at integer frame indices.  As shown by Takens \cite{takens1981detecting}, a sliding window of dimension $2m+1$ of even a single generic observation function (time series) of a dynamical system of intrinsic dimension $m$ is sufficient to reconstruct a topological embedding of the underlying trajectory in the original state space.  For periodic signals, this state space is a torus, and $d$ should be twice the number of harmonics present to injectively reconstruct loops on that torus \cite{perea2015sliding}.  The same is true of videos in general \cite{tralie2017quasi}.  Furthermore, the sliding window length $(\text{dim}-1) \tau$, or simply $\text{dim}-1$ in our case, maximizes the roundness of the embeddings when $\text{dim}-1 = k T$ for some integer $k$, where $T$ is the fundamental period.  Since we will use topological data analysis (TDA) to find the scale of the loop, and since TDA measures roundness, it is important that we tune our window size to be an integer multiple of the period.  To do this, we perform 1D ISOMAP \cite{tenenbaum2000global} on the original frames of the video to generate a 1D ``surrogate signal,'' and we perform an autocorrelation-based fundamental frequency estimation \cite{Mcleod05asmarter} on this signal to estimate the period and, hence, the appropriate dim.  This is similar to the diffusion maps + autocorrelation approach in \cite{tralie2017quasi} for analyzing periodic videos.  This estimation often returns integer multiples of the fundamental period, which is fine for our purposes.

@@

Since sliding window embeddings of periodic videos lie on a topological loop (a manifold which is an ``intrinsic circle''), this means that, at an appropriate scale, a nearest neighbors graph built on the sliding window point cloud will approximately be a circle graph.  To parameterize this graph, we turn to the graph Laplacian \cite{chung1997spectral}.  For a graph $G$ with $N$ vertices and an $N \times N$ symmetric adjacency matrix $A$, the graph Laplacian is defined as
\begin{equation}
L = D-A
\end{equation}
where $D_{ii} = \sum_{j = 1}^N A_{ij}, D_{i \neq j} = 0$ is the diagonal matrix.  Now let $d_{ij}: V \rightarrow \mathbb{R}^+$ be a nonnegative scalar function on the vertices of $G$.  Then the adjacency matrix correponding to the unweighted graph Laplacain, $A_{\epsilon}^u(d)$, is

@@

Alternatively, the weighted adjacency matrix is defined as $A_{\epsilon}^w(d)_{ij} = \exp(-d_{ij}^2/\epsilon)$ \cite{belkin2003laplacian}, but everything else is the same.  Now suppose we have performed a sliding window $SW_{\text{dim}}[X[n]]$ for a video $X$, where $X[n] = X[n+kT]$ for some integers $l, k$ so that there are $kT$ points in the sliding window (that is, $SW_{\text{dim}}[X[i]]$ repeats itself exactly every $T$ frames $k$ times total, for a total of $N = kT$ windows), and that $d_{ij} = ||SW_{\text{dim}}[X[i]] - SW_{\text{dim}}[X[j]]||_2$.  In this case, for any $\epsilon > 0$ and $|i - j| = kT$ for some integer $k$, $A^u_{\epsilon}(d)_{ij} = 1$.  Let us further suppose it is possible to choose an $\epsilon$ so that each window is also connected to the window directly before and after it in time, as well as the windows directly before and after its repetitions, but none others; that is, the unweighted adjacency matrix is a symmetric, {\em circulant matrix}, with the first row $A_{\epsilon}^u(d)_{1j} = 1$ if $j = 1, N-1,$ or $lT, 1 \leq l \leq (k-1)$.  Thus, the Laplacian is also circulant with first row

>>>>

\graphicspath{{Figures/}}
\newcommand{\argmin}{\operatornamewithlimits{argmin}}
\newcommand{\argmax}{\operatornamewithlimits{argmax}}
\newcommand{\matt}[1]{{\color{purple}[MSB: #1]}}
\newcommand{\mb}{\mathbf}

@@

\textbf{Preprocessing.} The input to our method is a 3-channel video of $N$ frames, each frame of resolution $W \times H$. We treat each frame as a point in Euclidean space of dimension $\mathbb{R}^{3 \cdot W \cdot H}$, and denote the set of all videos by the sequence $\mb{X}(t) \in \mathbb{R}^{(3 \cdot W \cdot H)}$ for integer-valued $t \in [1,N]$. The steps of our approach outlined in Figure~\ref{fig:ConceptFigure}(a-c), however, do not rely on the {\em extrinsic} geometry of the set of points $\mb{X}(t)$, only its set of {\em pairwise distances}. Thus, we can project all of the points into a lower-dimensional space and apply the set of techniques to the projection, assuming that pairwise distances are preserved. Since $N \ll (3 \cdot W \cdot H)$, the frames can be projected into a $N-1$-dimensional space such that all distances are preserved. Denoting this projection by $\mb{P} \in \mathbb{R}^{(N-1) \times (3 \cdot W \cdot H)}$, our method for steps (a)-(c) operates on $\mb{Z}(t) = \mb{P} \mb{X}(t)$.

@@

The sliding window video embedding \cite{cao1998dynamics,traliehigh,tralie2017quasi} for $\mb{Z}$ can defined as
\begin{equation}
	SW_{D}[\mb{Z}(t)] = \left[ \begin{array}{c} \mb{Z}(t) \\ \mb{Z}(t + 1) \\ \vdots \\ \mb{Z}(t + (D-1))  \end{array} \right] \in \mathbb{R}^{(N-1)\cdot(D)},
\end{equation}
where D is the dimension of the embedding. As shown by Takens~\cite{takens1981detecting}, a sliding window of dimension $2m+1$ of even a single generic observation function of a dynamical system of intrinsic dimension $m$ is sufficient to reconstruct a topological embedding of the underlying trajectory in the original state space.  For periodic signals, this state space is a torus, and $d$ should be twice the number of harmonics present to injectively reconstruct loops on that torus \cite{perea2015sliding}.  The same is true of videos in general \cite{tralie2017quasi}.  Furthermore, the sliding window length $(D-1)$ maximizes the roundness of the embeddings when $(D-1) = k T$ for some integer $k$, where $T$ is the fundamental period.  Since we will use topological data analysis (TDA) to find the scale of the loop, and since TDA measures roundness, it is important that we tune our window size to be an integer multiple of the period.  To do this, we perform 1D ISOMAP \cite{tenenbaum2000global} on the original frames of the video to generate a 1D surrogate signal, and we perform an autocorrelation-based fundamental frequency estimation \cite{Mcleod05asmarter} on this signal to estimate the period and, hence, the appropriate $D$.  This is similar to the diffusion maps + autocorrelation approach in \cite{tralie2017quasi} for analyzing periodic videos.  This estimation often returns integer multiples of the fundamental period, which is fine for our purposes.
% MSB: could probably remove the diffusion maps + autocorrelation approach sentence if we need space

@@

Since sliding window embeddings of periodic videos lie on a topological loop, this means that, at an appropriate scale, a nearest neighbors graph built on the sliding window point cloud will approximately be a graph representing a circle.  To parameterize this graph, we turn to the graph Laplacian \cite{chung1997spectral}. To define the graph Laplacian, we first construct an adjacency matrix $\mb{A}$ in which the entry at $\mb{A}_{ij}$ represents the similarity between windows at times $i$ and $j$: $\mb{A}_{ij} = exp(-||SW_{D}[\mb{Z}(i)] - SW_{D}[\mb{Z}(j)]||_2/\epsilon)$. The graph Laplacian then follows as:
\begin{equation}
	\mb{L} = \mb{D}-\mb{A}
\end{equation}
where $\mb{D}_{ii} = \sum_{j = 1}^N \mb{A}_{ij}, \mb{D}_{i \neq j} = 0$ is the diagonal matrix.  Now let $d_{ij}: V \rightarrow \mathbb{R}^+$ be a nonnegative scalar function on the vertices of $G$.  Then the adjacency matrix correponding to the unweighted graph Laplacain, $A_{\epsilon}^u(d)$, is

@@

Alternatively, the weighted adjacency matrix is defined as $A_{\epsilon}^w(d)_{ij} = \exp(-d_{ij}^2/\epsilon)$ \cite{belkin2003laplacian}, but everything else is the same.  Now suppose we have performed a sliding window $SW_{D}[X[n]]$ for a video $X$, where $X[n] = X[n+kT]$ for some integers $l, k$ so that there are $kT$ points in the sliding window (that is, $SW_{D}[X[i]]$ repeats itself exactly every $T$ frames $k$ times total, for a total of $N = kT$ windows), and that $d_{ij} = ||SW_{D}[X[i]] - SW_{D}[X[j]]||_2$.  In this case, for any $\epsilon > 0$ and $|i - j| = kT$ for some integer $k$, $A^u_{\epsilon}(d)_{ij} = 1$.  Let us further suppose it is possible to choose an $\epsilon$ so that each window is also connected to the window directly before and after it in time, as well as the windows directly before and after its repetitions, but none others; that is, the unweighted adjacency matrix is a symmetric, {\em circulant matrix}, with the first row $A_{\epsilon}^u(d)_{1j} = 1$ if $j = 1, N-1,$ or $lT, 1 \leq l \leq (k-1)$.  Thus, the Laplacian is also circulant with first row
